---
title: "Dev - frame class"
echo: false
format: 
  html:
    toc: true
    html-math-method: mathjax
    fig-width: 8
    fig-height: 7
vignette: >
  %\VignetteIndexEntry{Dev - frame class}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| echo: false
#| message: false

# library(officer.pptx)
devtools::load_all()
```

The `frame` class is a helper to fit an image to a placeholder. 

Two functions do the fitting:

- `frame_scale_to_target`: scale source to target frame. Scaling only, no change to  position (top, left)
- `frame_fit_to_target`: additional changes (translation, additional scaling etc.)

The following images show the behaviour for different settings visually.


## `frame_scale_to_target()`

To fit a source to a target frame, several `fit` modes exist:

- `none`: do not scale to target 
- `width`: scale source frame to match target's width
- `height`: scale source frame to match target's height
- `inside`: scale source frame so it fits inside target
- `outside`: scale source frame so it fits around target
- `fill`: scale to fill up the target frame (aspect ration not preserved)

Note that `inside` and `outside` either match `width` or `height`. Which depends on the frames.

See results for each fit mode.

```{r}
#| fig-height: 6

layout(matrix(1:6, ncol = 3, byrow = TRUE))

xlim <- c(0, 1.5)
ylim <- c(-1.5, 0)

f1 <- frame(0, 0, 1.5, 1.3)
f2 <- frame(0, 0, .5, 1)
f1_new <- frame_scale_to_target(f1, f2, fit = "none")
frames_draw(f1_new, f2, title = "original = none", xlim = xlim, ylim = ylim)
f1_new <- frame_scale_to_target(f1, f2, fit = "width")
frames_draw(f1_new, f2, title = "width", xlim = xlim, ylim = ylim)
f1_new <- frame_scale_to_target(f1, f2, fit = "h")
frames_draw(f1_new, f2, title = "height", xlim = xlim, ylim = ylim)
f1_new <- frame_scale_to_target(f1, f2, fit = "fill")
frames_draw(f1_new, f2, title = "fill", xlim = xlim, ylim = ylim)
f1_new <- frame_scale_to_target(f1, f2, fit = "inside")
frames_draw(f1_new, f2, title = "inside", xlim = xlim, ylim = ylim)
f1_new <- frame_scale_to_target(f1, f2, fit = "outside")
frames_draw(f1_new, f2, title = "outside", xlim = xlim, ylim = ylim)
```

Again, but with different original frame sizes.

```{r}
#| fig-height: 6

layout(matrix(1:6, ncol = 3, byrow = TRUE))

f1 <- frame(0, 0, 1.5, 1.3)
f2 <- frame(0, 0, 1, .5)
f1_new <- frame_scale_to_target(f1, f2, fit = "none")
frames_draw(f1_new, f2, title = "original = none", xlim = xlim, ylim = ylim)
f1_new <- frame_scale_to_target(f1, f2, fit = "width")
frames_draw(f1_new, f2, title = "width", xlim = xlim, ylim = ylim)
f1_new <- frame_scale_to_target(f1, f2, fit = "height")
frames_draw(f1_new, f2, title = "height", xlim = xlim, ylim = ylim)
f1_new <- frame_scale_to_target(f1, f2, fit = "fill")
frames_draw(f1_new, f2, title = "fill", xlim = xlim, ylim = ylim)
f1_new <- frame_scale_to_target(f1, f2, fit = "inside")
frames_draw(f1_new, f2, title = "inside", xlim = xlim, ylim = ylim)
f1_new <- frame_scale_to_target(f1, f2, fit = "outside")
frames_draw(f1_new, f2, title = "outside", xlim = xlim, ylim = ylim)
```

## `frame_fit_to_target()`

Order of operations (same as order of args):

1. Adjust source size to target frame
2. Scale by factor
3. Justify horizontal and vertical position within target
4. Add x/y-offset (extra space) as multiple of scaled image width/height

Justification with `fit = "none"` when target bigger or smaller than source frame.

```{r}
# center over target (h_hjust = .5, v_just = .5)  -----
layout(matrix(1:4, ncol = 2, byrow = TRUE))
xlim <- c(0, 1.5)
ylim <- c(-1.5, 1)
f1 <- frame(0, 0, 1.5, 1.3)
f2 <- frame(0, 0, .5, 1)

frames_draw(f1, f2, title = "original", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", v_just = .5, h_just = .5)
frames_draw(f1_new, f2, title = "fit: none, v_just = .5, h_just = .5", xlim = xlim, ylim = ylim)

frames_draw(f2, f1, title = "original", xlim = xlim, ylim = ylim)
f2_new <- frame_fit_to_target(f2, f1, fit = "none", v_just = .5, h_just = .5)
frames_draw(f2_new, f1, title = "fit: none, v_just = .5, h_just = .5", xlim = xlim, ylim = ylim)
```

Varying with `fit = "none"` varying `h_just` and `v_just` when target is bigger than source frame.

```{r}
# h_just, v_just: target bigger than source ----
layout(matrix(1:4, ncol = 2, byrow = TRUE))
xlim <- c(0, 1.5)
ylim <- c(-1.5, 1)
f1 <- frame(.2, 0, 1.5, 1.3)
f2 <- frame(0, 0, .5, 1)

frames_draw(f2, f1, title = "original", xlim = xlim, ylim = ylim)
f2_new <- frame_fit_to_target(f2, f1, fit = "none", v_just = .5, h_just = .5)
frames_draw(f2_new, f1, title = "fit: none, v_just = .5, h_just = .5", xlim = xlim, ylim = ylim)
f2_new <- frame_fit_to_target(f2, f1, fit = "none", v_just = 1, h_just = 0)
frames_draw(f2_new, f1, title = "fit: none, v_just = 1, h_just = .0", xlim = xlim, ylim = ylim)
f2_new <- frame_fit_to_target(f2, f1, fit = "none", v_just = 0, h_just = 1)
frames_draw(f2_new, f1, title = "fit: none, v_just = 0, h_just = 1", xlim = xlim, ylim = ylim)
```

Varying with `fit = "none"` varying `h_just` and `v_just` when source is bigger than target frame.

```{r}
# h_just, v_just: source bigger than target ----
layout(matrix(1:4, ncol = 2, byrow = TRUE))
xlim <- c(0, 1.5)
ylim <- c(-1.5, 1)
f2 <- frame(.2, 0, 1.5, 1.3)
f1 <- frame(0, 0, .5, 1)

frames_draw(f2, f1, title = "original", xlim = xlim, ylim = ylim)
f2_new <- frame_fit_to_target(f2, f1, fit = "none", v_just = .5, h_just = .5)
frames_draw(f2_new, f1, title = "fit: none, v_just = .5, h_just = .5", xlim = xlim, ylim = ylim)
f2_new <- frame_fit_to_target(f2, f1, fit = "none", v_just = 1, h_just = 0)
frames_draw(f2_new, f1, title = "fit: none, v_just = 1, h_just = 0", xlim = xlim, ylim = ylim)
f2_new <- frame_fit_to_target(f2, f1, fit = "none", v_just = 0, h_just = 1)
frames_draw(f2_new, f1, title = "fit: none, v_just = 0, h_just = 1", xlim = xlim, ylim = ylim)
```

Varying `h_just`, `x_offset`, and `offset_mode`. All combinations yield same results pre row.

```{r}
#| fig-height: 6
#| 
# x_offset, y_offset: source bigger than target ----
layout(matrix(1:8, ncol = 4, byrow = TRUE))
xlim <- c(0, 2.5)
ylim <- c(-1.5, .5)
f2 <- frame(.2, 0, 1.5, 1.3)
f1 <- frame(0, 0, .5, 1)

# three offset modes with same results
f1_new <- frame_fit_to_target(f1, f2, fit = "none")
frames_draw(f1_new, f2, title = "original", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", x_offset = 1, h_just = 1, offset_mode = "source")
frames_draw(f1_new, f2, title = "x_offset = 1 \nh_just = 1 \noffset_mode = source", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", x_offset = 1, h_just = 0, offset_mode = "target")
frames_draw(f1_new, f2, title = "x_offset = 1 \nh_just = 0 \noffset_mode = target", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", x_offset = f2$width, h_just = 0, offset_mode = "units")
frames_draw(f1_new, f2, title = "x_offset = target$width \nh_just = 0 \noffset_mode = units", xlim = xlim, ylim = ylim)

# three offset modes with same results
f1_new <- frame_fit_to_target(f1, f2, fit = "none")
frames_draw(f1_new, f2, title = "original", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", x_offset = -1, h_just = 0, offset_mode = "source")
frames_draw(f1_new, f2, title = "x_offset = -1\nh_just = 0\noffset_mode = source", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", x_offset = -1, h_just = 1, offset_mode = "target")
frames_draw(f1_new, f2, title = "x_offset = -1\nh_just = 1\noffset_mode = target", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", x_offset = - f2$width, h_just = 1, offset_mode = "units")
frames_draw(f1_new, f2, title = "x_offset = -target$width \nh_just = 1 \noffset_mode = units", xlim = xlim, ylim = ylim)
```

Varying `v_just`, `y_offset`, and `offset_mode`. All combiantion yield same results per row.

```{r}
#| fig-height: 6
#| 
layout(matrix(1:8, ncol = 4, byrow = TRUE))

# three offset modes with same results
f1_new <- frame_fit_to_target(f1, f2, fit = "none")
frames_draw(f1_new, f2, title = "original", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", y_offset = 1, v_just = 1, offset_mode = "source")
frames_draw(f1_new, f2, title = "y_offset = 1 \nv_just = 1 \noffset_mode = source", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", y_offset = 1, v_just = 0, offset_mode = "target")
frames_draw(f1_new, f2, title = "y_offset = 1 \nv_just = 0 \noffset_mode = target", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", y_offset = f2$height, v_just = 0, offset_mode = "units")
frames_draw(f1_new, f2, title = "y_offset = target$height \nv_just = 0 \noffset_mode = units", xlim = xlim, ylim = ylim)

# three offset modes with same results
f1_new <- frame_fit_to_target(f1, f2, fit = "none")
frames_draw(f1_new, f2, title = "original", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", y_offset = -1, v_just = 0, offset_mode = "source")
frames_draw(f1_new, f2, title = "y_offset = -1 \nv_just = 0 \noffset_mode = source", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", y_offset = -1, v_just = 1, offset_mode = "target")
frames_draw(f1_new, f2, title = "y_offset = -1 \nv_just = 1 \noffset_mode = target", xlim = xlim, ylim = ylim)
f1_new <- frame_fit_to_target(f1, f2, fit = "none", y_offset = -f2$height, v_just = 1, offset_mode = "units")
frames_draw(f1_new, f2, title = "y_offset = -target$height \nv_just = 1 \noffset_mode = units", xlim = xlim, ylim = ylim)
```
