---
title: "Dev"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dev}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(officer.pptx)
x <- read_pptx()
```


## Placeholders

Different from `officer::ph_with`, `phs_with` fills any number of placeholders by passing the placeholders short-from syntax and an object as a key-value pair.

```{r}
x <- x |>
  add_slide("Two Content") |>
  phs_with(
    `Title 1` = "A title", dt = "Jan. 26, 2025",
    `body[2]` = "Body 2", left = "Left side", `6` = "Footer"
  )
```

`add_slide` is patched with additional dots to fill placeholders directly when adding a slide.

```{r}
x <- x |> add_slide("Two Content",
  `Title 1` = "A title", dt = "Jan. 26, 2025",
  `body[2]` = "Body 2", left = "Left side", `6` = "Footer"
)
```

## Replace text

Replace text in a shape while keeping the text formats intact.

```{r}
file_in <- example_pptx("text_replace")
file_out <- tempfile(fileext = ".pptx")
x <- pptx_read(file_in)
x <- text_replace(x, "@" = ">>>", "{1}" = ":)")
# print(x, target = file_out)
# browseURL(file_out)
```

## Add image

Add image, with arbitrary shape acting as a placeholder.

```{r add-images}
file <- example_pptx("image_insert")

x <- read_pptx(file)

x <- image_insert(x, example_image("dog_1"), pattern = "{2}", h_just = 0)
x <- image_insert(x, example_image("dog_2"), pattern = "{3}", h_just = .5)
x <- image_insert(x, example_image("dog_2"), pattern = "{4}", h_just = .5)

# file_out <- tempfile(fileext = ".pptx")
# print(x, file_out)
# browseURL(file_out)
```


## Preview

Create a preview of a slide.

```{r slide-preview}
slide_preview(x, 1)
slide_preview(x, 2)
```



## Shapes

We want to be able to modify existing and add new shapes. The shape objects have a lot of properties and method. 
For our purpose, we will only implement a small subset. Also, shapes can have hierarchical structures, such as group shapes
etc. For our template engine, we will not model these hierarchies but treat all shapes the same, no matter of its location.

For reporting, the most common case will be to copy a template slide and populate or modify the shapes on the slide, 
that serve as placeholders.

Our use case requires:

- Add shape at specified position
- Delete shape
- Modify shape's background, color, border, size, text, visibility etc.
- (Maybe: group shapes)

We use a `<shapes>` class containing `<shape>` objects, for example, from a slide. Each shape can be modified.

A slide can be plotted, to you can see the size and position as well as its approximate content. 

```{r}
file <- example_pptx("image_insert")
x <- read_pptx(file)
# pptx_shapes_info(x)
```

There are some shape-like structures (similar properties):

- Shape: either be a preset or a custom geometry
- Picture
- ConnectorShape
- GroupShape

### ShapeTree

A ShapeTree `<p:spTree>` contains all shapes, pictures, graphicframes and groupshapes from a slide.
The order of the xml node in the shapetree determine the z-order. 

For our purposes, it will mostly suffice to treat all shapes the same ignoring any hierarchical orders caused by group shapes.
For replacing text this will be steightforward. However, when detrminig the x,y coordinates, one needs to take the 
hierarchical structure into account: A child shape’s `<a:off>` is not absolute. It’s only relative to the group shape's `chOff`. 
Calculate as follows: `absoluteX = group.off.x + (child.off.x - group.chOff.x)`

### GroupShape

A groupshape serves as a new frame of reference for the contained shapes, with the x,y coord now being recorded relative to the group. 
Within a GroupShape, the z-order is determined by the node order as well.



### Shape

The dotnet [Shape class](https://learn.microsoft.com/en-us/dotnet/api/documentformat.openxml.drawing.shape?view=openxml-3.0.1)
has these subelements:

- nvSpPr (Non-Visual Properties for a Shape)	§20.1.2.2.29
- spPr (Shape Properties)	§20.1.2.2.35
- style (Shape Style)	§20.1.2.2.37
- txSp (Text Shape)	§20.1.2.2.41
- (extLst (Extension List)	§20.1.2.2.15)

## ADRs

- In the first step it will suffice to ...


## Todos

- [ ] Build a class for Shape, Picture, GraphicFrame, ShapeTree
- [ ] Refactor and add inheritance if useful
- [ ] Return correct coordinates for shapes inside group shapes
- [ ] Basis translation PoC of commonmark syntax to `<p:txBody>`


## Ideas

Usually, we have a complicated custom template. Different header colors and icons for different topics.
Several types of templates, depending on the question set.

Probably, this could be simplified to some VAC (visuals-as-code) plus SAC (slides-as-code) using YAML, TOML etc.

```
- visual:
  - name : viz_1
  - type: barchart_simple
  - vars: gender
  - filters: 
    - org: 1
...

- slide: 
  - layout: "skala mit items",
  - title.fill : "#333333",
  - title.icon: "path/to/img/",
  - title.text : "Team Cooperation",
  - subtitle.fill : "#333333",
  - subtitle.text : "Your teams vs other teams"
  - my_chart.visual: viz_1
...

```

## Learnings

Is it possible to 

```{r eval=FALSE}
file_in <- "../dev_local/xx_fill_phs/template.pptx"
file_out <- "../dev_local/xx_fill_phs/template_FILLED.pptx"

# x <- read_pptx()
# print(x, "../dev_local/xx_fill_phs/template.pptx")

x <- read_pptx(file_in)
plot_layout_properties(x, "l1")
x <- x |> add_slide("l1", "title" = "Title Text", "body" = "PH 1")

print(x, file_out)
```


