---
title: "Inserting Images with img()"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: false
vignette: >
  %\VignetteIndexEntry{Inserting Images with img()}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(officer.pptx)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

render_miniature <- function(name) {
  source_path <- system.file(
    "ext/vignettes/img-function-thumbnails",
    name,
    package = "officer.pptx"
  )
  local_dir <- "img-function-thumbnails"
  dir.create(local_dir, showWarnings = FALSE)
  local_path <- file.path(local_dir, name)
  if (!file.exists(local_path)) {
    file.copy(source_path, local_path)
  }
  knitr::include_graphics(local_path)
}
```

## Introduction

The `img()` function is a powerful replacement for `officer::external_img()` that makes inserting images into PowerPoint placeholders much easier. The key improvements are:
 
- **Automatic size detection**: No need to manually specify image dimensions
- **Smart positioning**: Images are automatically scaled and centered within placeholders
- **Flexible alignment**: Control horizontal and vertical positioning with justification parameters
- **Styling options**: Add borders, backgrounds, and rotation
- **Multiple image formats**: Support for PNG, JPEG, GIF, TIFF, SVG, PDF, EMF, and WMF

## The Problem with `external_img()`
 
When using `officer::external_img()`, you typically need to specify the image dimensions manually. If you don't, the image may not fit properly within the placeholder. The `img()` function solves this by automatically detecting image dimensions and fitting the image inside the placeholder while preserving the aspect ratio.

## Supported Image Formats

The `img()` function supports automatic size detection for the following formats:

| Format | Extension | Size Detection |
|--------|-----------|----------------|
| PNG | `.png` | via `magick` |
| JPEG | `.jpg`, `.jpeg` | via `magick` |
| GIF | `.gif` | via `magick` |
| TIFF | `.tiff` | via `magick` |
| SVG | `.svg` | via `magick` |
| PDF | `.pdf` | via `pdftools` |
| EMF | `.emf` | native (header parsing) |
| WMF | `.wmf` | native (placeable header) |

::: {.callout-note}
WMF files require an Aldus Placeable Metafile header for automatic size detection. If your WMF file doesn't have this header, you can still use it by manually specifying `.width` and `.height`.
:::

## Basic Usage

The simplest way to use `img()` is to just pass the image path. The image will be automatically scaled to fit inside the placeholder and centered.

::: {.callout-tip}
Throughout this vignette, the dashed gray border indicates the placeholder boundary. It is added using `ph_visualize()` via the helper function `ph_frame()` defined below.
:::

```{r}
#| echo: true

# Helper to add a dashed gray placeholder frame
ph_frame <- function(x, ph = "body") {
  ph_visualize(x, ph = ph, line = sp_line(color = "gray50", lwd = 1.5, lty = "dash"))
}

# Create a sample image
file_png <- tempfile(fileext = ".png")
png(file_png, width = 500, height = 400)
plot(1:10, col = "steelblue", pch = 16, cex = 2, 
     xlab = "X", ylab = "Y", main = "Sample Plot")
box("figure", col = "gray60", lwd = 3)
dev.off()

# Insert into a slide - that's it!
file_pptx <- tempfile(fileext = ".pptx")
x <- read_pptx() |>
  layout_default("Title and Content") |>
  add_slide(
    title = "Basic Usage - Automatic Fit",
    body = img(file_png)
  ) |>
  ph_frame()
print(x, file_pptx)
```

```{r}
#| echo: false
#| fig-width: 8
#| fig-height: 4.5

render_miniature("thumb-01.png")
```

Compare this to the traditional approach with `external_img()` where you would need to specify dimensions manually or deal with improper sizing.

## Function Parameters

The `img()` function offers many parameters to control image placement and styling:

```r
img(
 src,                    # Image file path
 fit = "inside",         # How to fit: "inside", "outside", "width", "height", "fill", "none"
 scale = 1,              # Scale factor (0-1 for smaller, >1 for larger)
 h_just = 0.5,           # Horizontal justification (0=left, 0.5=center, 1=right)
 v_just = 0.5,           # Vertical justification (0=bottom, 0.5=center, 1=top)
 x_offset = 0,           # Horizontal offset
 y_offset = 0,           # Vertical offset
 offset_mode = "source", # How to interpret offsets: "source", "target", "units"
 rotation = 0,           # Rotation in degrees (anti-clockwise)
 background = "transparent",  # Background color
 line = sp_line(),       # Border specification
 alt = "",               # Alternative text for accessibility
 unit = "in",            # Unit for manual dimensions: "in", "cm", "mm"
 .left = NULL,           # Override left position
 .top = NULL,            # Override top position
 .width = NULL,          # Override width
 .height = NULL          # Override height
)
```

## Scaling and Fitting

### The `scale` Parameter

The `scale` parameter allows you to resize the image after fitting:

```{r}
#| echo: true

file_pptx <- tempfile(fileext = ".pptx")
x <- read_pptx() |>
  layout_default("Title and Content") |>
  add_slide(
    title = "scale = 1 (default, full size)",
    body = img(file_png, scale = 1)
  ) |>
  ph_frame() |>
  add_slide(
    title = "scale = 0.7 (70% size)",
    body = img(file_png, scale = 0.7)
  ) |>
  ph_frame() |>
  add_slide(
    title = "scale = 0.5 (50% size)",
    body = img(file_png, scale = 0.5)
  ) |>
  ph_frame()
print(x, file_pptx)
```

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 3

render_miniature("thumb-02.png")
```

### The `fit` Parameter

The `fit` parameter controls how the image is scaled relative to the placeholder:

- **`"inside"`** (default): Scale image to fit entirely inside the placeholder
- **`"outside"`**: Scale image to completely cover the placeholder
- **`"width"`**: Match the placeholder width exactly
- **`"height"`**: Match the placeholder height exactly
- **`"fill"`**: Stretch to fill (may distort aspect ratio)
- **`"none"`**: Use original image size

The dashed gray border in the examples below shows the placeholder boundary using `ph_visualize()`:

```{r}
#| echo: true

file_pptx <- tempfile(fileext = ".pptx")
x <- read_pptx() |>
  layout_default("Title and Content") |>
  add_slide(title = "fit = 'inside' (default)") |>
  ph_with(img(file_png, fit = "inside"), location = "body") |>
  ph_frame() |>
  add_slide(title = "fit = 'width'") |>
  ph_with(img(file_png, fit = "width"), location = "body") |>
  ph_frame() |>
  add_slide(title = "fit = 'height'") |>
  ph_with(img(file_png, fit = "height"), location = "body") |>
  ph_frame() |>
  add_slide(title = "fit = 'fill' (stretches)") |>
  ph_with(img(file_png, fit = "fill"), location = "body") |>
  ph_frame()
print(x, file_pptx)
```

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 5

render_miniature("thumb-03.png")
```

## Positioning and Alignment

### Justification with `h_just` and `v_just`

Control where the image is positioned within the placeholder:

- `h_just`: Horizontal position (0 = left, 0.5 = center, 1 = right)
- `v_just`: Vertical position (0 = bottom, 0.5 = center, 1 = top)

```{r}
#| echo: true

file_pptx <- tempfile(fileext = ".pptx")
x <- read_pptx() |>
  layout_default("Title and Content") |>
  add_slide(
    title = "Top-Left (h_just=0, v_just=1)",
    body = img(file_png, scale = 0.6, h_just = 0, v_just = 1)
  ) |>
  ph_frame() |>
  add_slide(
    title = "Center (default)",
    body = img(file_png, scale = 0.6, h_just = 0.5, v_just = 0.5)
  ) |>
  ph_frame() |>
  add_slide(
    title = "Bottom-Right (h_just=1, v_just=0)",
    body = img(file_png, scale = 0.6, h_just = 1, v_just = 0)
  ) |>
  ph_frame()
print(x, file_pptx)
```

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 3

render_miniature("thumb-04.png")
```

### Offsets with `x_offset` and `y_offset`

Add extra spacing to fine-tune positioning. By default, offsets are multiples of the scaled image size:

```{r}
#| echo: true

file_pptx <- tempfile(fileext = ".pptx")
x <- read_pptx() |>
  layout_default("Title and Content") |>
  add_slide(title = "Using x_offset for side-by-side images") |>
  phs_with(
    body = img(file_png, scale = 0.4, x_offset = -1.1),
    body = img(file_png, scale = 0.4, x_offset = 0),
    body = img(file_png, scale = 0.4, x_offset = 1.1)
  ) |>
  ph_frame()
print(x, file_pptx)
```

```{r}
#| echo: false
#| fig-width: 8
#| fig-height: 4.5

render_miniature("thumb-05.png")
```

The `offset_mode` parameter controls how offsets are interpreted:

- **`"source"`** (default): Offset as multiple of the scaled image's width/height
- **`"target"`**: Offset as multiple of the placeholder's width/height
- **`"units"`**: Offset in raw units (inches by default, see `unit` parameter)

## Styling Options

### Rotation

Rotate the image anti-clockwise:

```{r}
#| echo: true

file_pptx <- tempfile(fileext = ".pptx")
x <- read_pptx() |>
  layout_default("Title and Content") |>
  add_slide(title = "rotation = 0 (default)", body = img(file_png, scale = 0.7)) |>
  ph_frame() |>
  add_slide(title = "rotation = 15", body = img(file_png, scale = 0.7, rotation = 15)) |>
  ph_frame() |>
  add_slide(title = "rotation = 45", body = img(file_png, scale = 0.7, rotation = 45)) |>
  ph_frame()
print(x, file_pptx)
```

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 3

render_miniature("thumb-06.png")
```

### Background Color

Add a background color behind the image (useful for images with transparency):

```{r}
#| echo: true

# Create image with transparent background
file_png_transp <- tempfile(fileext = ".png")
png(file_png_transp, width = 600, height = 400, bg = "transparent")
plot(1:10, col = "steelblue", pch = 16, cex = 3,
     xlab = "X", ylab = "Y", main = "Transparent Background")
dev.off()

file_pptx <- tempfile(fileext = ".pptx")
x <- read_pptx() |>
  layout_default("Title and Content") |>
  add_slide(
    title = "No background (transparent)",
    body = img(file_png_transp, scale = 0.7)
  ) |>
  ph_frame() |>
  add_slide(
    title = "background = 'lightblue'",
    body = img(file_png_transp, scale = 0.7, background = "lightblue")
  ) |>
  ph_frame() |>
  add_slide(
    title = "background = '#FFE4E1' (misty rose)",
    body = img(file_png_transp, scale = 0.7, background = "#FFE4E1")
  ) |>
  ph_frame()
print(x, file_pptx)
```

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 3

render_miniature("thumb-07.png")
```

### Border / Line

Add a border around the image. You can use a simple color name or an `sp_line()` object for more control:

```{r}
#| echo: true

file_pptx <- tempfile(fileext = ".pptx")
x <- read_pptx() |>
  layout_default("Title and Content") |>
  add_slide(
    title = "Simple border: line = 'red'",
    body = img(file_png, scale = 0.7, line = "red")
  ) |>
  ph_frame() |>
  add_slide(
    title = "Thick dashed border",
    body = img(file_png, scale = 0.7, 
               line = sp_line(color = "darkblue", lwd = 3, lty = "dash"))
  ) |>
  ph_frame() |>
  add_slide(
    title = "Combined: border + background",
    body = img(file_png, scale = 0.7, 
               line = sp_line(color = "darkgreen", lwd = 2),
               background = "#E8F5E9")
  ) |>
  ph_frame()
print(x, file_pptx)
```

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 3

render_miniature("thumb-08.png")
```

## Multiple Images on One Slide

You can add multiple images to the same placeholder using `phs_with()`:

```{r}
#| echo: true

file_pptx <- tempfile(fileext = ".pptx")
x <- read_pptx() |>
  layout_default("Title and Content") |>
  add_slide(title = "Multiple images with different positions") |>
  phs_with(
    body = img(file_png, scale = 0.45, h_just = 0, v_just = 1, line = "blue"),
    body = img(file_png, scale = 0.45, h_just = 1, v_just = 0, line = "red"),
    body = img(file_png, scale = 0.35, h_just = 0.5, v_just = 0.5, 
               rotation = 10, line = "green")
  ) |>
  ph_frame()
print(x, file_pptx)
```

```{r}
#| echo: false
#| fig-width: 8
#| fig-height: 4.5

render_miniature("thumb-09.png")
```

## Using `img()` with `ph_with()`

The `img()` function works seamlessly with `officer::ph_with()`. You can even override `img()` settings in the `ph_with()` call:

```{r}
#| echo: true

file_pptx <- tempfile(fileext = ".pptx")
x <- read_pptx() |>
  layout_default("Title and Content") |>
  add_slide(title = "Settings in img()") |>
  ph_with(
    value = img(file_png, scale = 0.7, line = "blue"),
    location = "body"
  ) |>
  ph_frame() |>
  add_slide(title = "Override settings in ph_with()") |>
  # Caveat: Does not work yet. Requires https://github.com/davidgohel/officer/issues/681 to be finished first.
  ph_with(
    value = img(file_png, scale = 0.7),
    location = "body",
    rotation = 20,
    background = "#FFF3E0",
    line = sp_line(color = "orange", lwd = 2)
  ) |>
  ph_frame()
print(x, file_pptx)
```

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 4.5

render_miniature("thumb-10.png")
```

## Manual Dimension Override

For cases where automatic size detection doesn't work (e.g., WMF files without placeable header) or when you want specific dimensions, use the `.left`, `.top`, `.width`, and `.height` parameters:

```{r}
#| eval: false

# Manually specify image dimensions (in inches by default)
add_slide(body = img(
  "image_without_size_info.wmf",
  .width = 4,
  .height = 3
))

# Use different units
add_slide(body = img(
  file_png,
  .width = 100,
  .height = 80,
  unit = "mm"
))

# Completely custom position and size
add_slide(body = img(
  file_png,
  .left = 1,
 .top = 1,
  .width = 3,
  .height = 2,
  unit = "in"
))
```

## Complete Example

Here's a complete example demonstrating various `img()` features:

```{r}
#| echo: true

# Create sample image
file_plot <- tempfile(fileext = ".png")
png(file_plot, width = 600, height = 400)
plot(mtcars$mpg, mtcars$hp, 
     col = "steelblue", pch = 19, cex = 1.5,
     xlab = "MPG", ylab = "Horsepower",
     main = "Motor Trend Car Data")
box("figure", col = "gray60", lwd = 3)
dev.off()

# Build presentation
file_pptx <- tempfile(fileext = ".pptx")
x <- read_pptx() |>
  layout_default("Title and Content") |>
  # Slide 1: Basic insertion
  add_slide(title = "Default Settings", body = img(file_plot)) |>
  ph_frame() |>
  # Slide 2: Scaled and positioned
  add_slide(
    title = "Scaled 70%, Top-Left",
    body = img(file_plot, scale = 0.7, h_just = 0, v_just = 1)
  ) |>
  ph_frame() |>
  # Slide 3: With styling
  add_slide(
    title = "With Border and Background",
    body = img(file_plot, scale = 0.8, background = "lightblue",
               line = sp_line(color = "gray40", lwd = 2))
  ) |>
  ph_frame() |>
  # Slide 4: Rotated
  add_slide(
    title = "Rotated 15 Degrees",
    body = img(file_plot, scale = 0.7, rotation = 15)
  ) |>
  ph_frame() |>
  # Slide 5: Multiple images
  add_slide(title = "Multiple Images") |>
  phs_with(
    body = img(file_plot, scale = 0.45, h_just = 0, v_just = 1, line = "steelblue"),
    body = img(file_plot, scale = 0.45, h_just = 1, v_just = 0, line = "coral", rotation = 8)
  ) |>
  ph_frame()

print(x, file_pptx)
```

```{r}
#| echo: false
#| fig-width: 10
#| fig-height: 6

render_miniature("thumb-11.png")
```

## Tips and Best Practices

1. **Use `fit = "inside"` (default)** for most cases to ensure the entire image is visible.

2. **Combine `scale` with justification** to position smaller images exactly where you want them.

3. **Use `sp_line()` for detailed border control** when you need specific line widths or dash patterns.

4. **Set `alt` text** for accessibility compliance in professional presentations.

5. **Use `offset_mode = "units"`** when you need precise pixel-perfect positioning.

6. **For WMF files**, ensure they have a placeable header, or provide manual dimensions.

## See Also

- `officer::external_img()` - The original function that `img()` replaces
- `officer::ph_with()` - For adding content to placeholders
- `officer::sp_line()` - For border/line specifications
- `phs_with()` - For adding multiple items to placeholders
