---
title: "Selecting Shapes with shape_select()"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
vignette: >
  %\VignetteIndexEntry{Selecting Shapes with shape_select()}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(officer.pptx)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Purpose

`shape_select()` selects existing top-level objects on PowerPoint slides. It is meant for template workflows where manually placed shapes act as markers for later processing.

A common pattern is to place a rectangle on a slide and put a visible marker such as `{image_1}` inside it. `shape_select()` retrieves those markers and returns their slide coordinates, metadata, and XML nodes. It does not decide what happens next.

```{r}
x <- read_pptx()
x <- add_slide(x, layout = "Blank")
x <- ph_with(
  x,
  "{image_1}",
  location = ph_location(
    left = 1, top = 1, width = 3, height = 1.5,
    newlabel = "Image marker"
  )
)

shape_select(x, text = "{image_1}")
```

## Current Feature Scope

The current implementation is intentionally selection-only.

Supported:

- Select top-level slide objects.
- Search one slide, several slides, or all slides.
- Match by visible text and Selection Pane name.
- Filter by XML ID, object kind, placeholder status, placeholder type, and hidden state.
- Return geometry in slide-level inches.
- Preserve visible presentation order after slide reordering.
- Keep the PowerPoint cursor unchanged.
- Return duplicate matches rather than forcing uniqueness.

Not supported yet:

- Recursively selecting shapes inside groups.
- Replacing, hiding, removing, or inserting content.
- Treating XML IDs as persistent identifiers.
- Resolving layout placeholders such as `ph_location_label()`; those are a separate officer concept.

## Matching

The `match` argument controls how `text` and `name` filters are interpreted.

| Mode | Behavior |
|------|----------|
| `"contains"` | Default. Fixed substring matching. |
| `"exact"` | Complete-value matching after trimming leading and trailing whitespace. |
| `"regex"` | Regular-expression matching. |

### Contains Matching

`contains` is the default because marker text is often embedded in helper text while designing templates.

```{r}
x <- add_slide(x, layout = "Blank")
x <- ph_with(
  x,
  "Insert {image_1} here",
  location = ph_location(
    left = 1, top = 1, width = 3, height = 1.5,
    newlabel = "Instruction marker"
  )
)

shape_select(x, text = "{image_1}")[, c("slide_idx", "name", "text")]
```

### Exact Matching

Use `match = "exact"` when the whole shape text should be the marker. Leading and trailing whitespace are ignored.

```{r}
shape_select(x, text = "{image_1}", match = "exact")[, c("slide_idx", "name", "text")]
```

### Regex Matching

Use `match = "regex"` for marker families.

```{r}
shape_select(
  x,
  text = "^\\s*\\{image_[0-9]+\\}\\s*$",
  match = "regex"
)[, c("slide_idx", "name", "text")]
```

## Filtering by Metadata

Different criteria are combined with AND. Multiple values within one criterion are alternatives.

```{r}
shape_select(
  x,
  text = "{image_1}",
  name = "marker"
)[, c("slide_idx", "shape_idx", "name", "text", "kind")]
```

Object kinds currently include:

| Kind | Meaning |
|------|---------|
| `"shape"` | Regular PowerPoint shape or text box. |
| `"picture"` | Picture object. |
| `"table"` | Table graphic frame. |
| `"chart"` | Chart graphic frame. |
| `"graphic_frame"` | Other graphic frame. |
| `"group"` | Top-level group. |
| `"connector"` | Connector shape. |

## Result Columns

`shape_select()` returns a tibble-like `pptx_shape_selection` object. The `node` list-column is kept in the object for downstream functions, but hidden by the print method.

Important columns:

| Column | Meaning |
|--------|---------|
| `slide_idx` | Visible presentation slide index. |
| `shape_idx` | Top-level shape-tree position, which approximates z-order. |
| `id` | Current slide-local XML shape ID. Not stable across save operations. |
| `name` | Selection Pane name. |
| `text` | Visible text. For groups this is `NA` in the first implementation. |
| `kind` | Object kind. |
| `placeholder` | Whether the object is an actual placeholder. |
| `ph_type` | Placeholder type, if present. |
| `hidden` | Whether `p:cNvPr/@hidden` is set to `"1"`. |
| `left`, `top`, `width`, `height` | Slide-level frame in inches. |
| `rotation` | Rotation in degrees, if present. |
| `description` | Alternative text description, if present. |
| `node` | XML node reference for downstream code. |

```{r}
targets <- shape_select(x, text = "{image_1}")
names(targets)
```

## Groups

The first implementation is top-level only. A group is returned as one row with `kind = "group"`; shapes inside the group are not listed or matched.

This keeps coordinates simple and reliable. Recursive group support would need separate handling for group-relative coordinates, nested z-order, rotation, and possible duplicate matches between a group and its children.

## Relation to officer Placeholders

`shape_select()` works on actual slide XML nodes. It is different from officer's `ph_location_*()` functions, which resolve layout placeholders.

Use `ph_location_*()` when you are creating new content from a layout. Use `shape_select()` when an existing slide already contains marker shapes that should be discovered.

## Typical Next Step

`shape_select()` intentionally stops at retrieval. A downstream function can decide whether to overlay, replace, hide, or remove the selected marker shapes.

```{r}
#| eval: false
targets <- shape_select(template, text = "{image_1}")

# Future consumers can use targets$left, targets$top, targets$width,
# targets$height, and targets$node.
```
