---
title: "Selecting and Updating Shapes"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: false
vignette: >
  %\VignetteIndexEntry{Selecting and Updating Shapes}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(officer.pptx)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

shape_thumbnail <- function(name) {
  vapply(name, function(one_name) {
    source_path <- system.file(
      "ext/vignettes/shape-select-thumbnails",
      one_name,
      package = "officer.pptx"
    )
    local_dir <- "shape-select-thumbnails"
    dir.create(local_dir, showWarnings = FALSE)
    local_path <- file.path(local_dir, one_name)
    if (!file.exists(local_path)) {
      file.copy(source_path, local_path)
    }
    local_path
  }, character(1))
}
```

## Purpose

`shape_select()` selects existing top-level objects on PowerPoint slides. `shape_update()` can then modify the selected objects in place. Together they support template workflows where manually placed shapes act as markers for later processing.

A common pattern is to place a rectangle on a slide and put a visible marker such as `{image_1}` inside it. `shape_select()` retrieves those markers and returns their slide coordinates, metadata, and XML nodes.

```{r}
x <- read_pptx()
x <- add_slide(x, layout = "Blank")
x <- ph_with(
  x,
  "{image_1}",
  location = ph_location(
    left = 1, top = 1, width = 3, height = 1.5,
    newlabel = "Image marker"
  )
)

shape_select(x, text = "{image_1}")
```

## Current Feature Scope

The current implementation covers top-level shape selection and focused updates of selected shapes.

Supported:

- Select top-level slide objects.
- Search one slide, several slides, or all slides.
- Match by visible text and Selection Pane name.
- Filter by XML ID, object kind, placeholder status, placeholder type, and hidden state.
- Return geometry in slide-level inches.
- Preserve visible presentation order after slide reordering.
- Keep the PowerPoint cursor unchanged.
- Return duplicate matches rather than forcing uniqueness.
- Modify selected shapes with `shape_update()`.
- Update ordinary non-placeholder shapes; placeholder status is metadata, not a prerequisite.

Not supported yet:

- Recursively selecting shapes inside groups.
- Removing shapes, inserting new shapes, or replacing pictures.
- Detailed run-level text formatting.
- Treating XML IDs as persistent identifiers.
- Resolving layout placeholders such as `ph_location_label()`; those are a separate officer concept.

## Matching

The `match` argument controls how `text` and `name` filters are interpreted.

| Mode | Behavior |
|------|----------|
| `"contains"` | Default. Fixed substring matching. |
| `"exact"` | Complete-value matching after trimming leading and trailing whitespace. |
| `"regex"` | Regular-expression matching. |

### Contains Matching

`contains` is the default because marker text is often embedded in helper text while designing templates.

```{r}
x <- add_slide(x, layout = "Blank")
x <- ph_with(
  x,
  "Insert {image_1} here",
  location = ph_location(
    left = 1, top = 1, width = 3, height = 1.5,
    newlabel = "Instruction marker"
  )
)

shape_select(x, text = "{image_1}")[, c("slide_idx", "name", "text")]
```

### Exact Matching

Use `match = "exact"` when the whole shape text should be the marker. Leading and trailing whitespace are ignored.

```{r}
shape_select(x, text = "{image_1}", match = "exact")[, c("slide_idx", "name", "text")]
```

### Regex Matching

Use `match = "regex"` for marker families.

```{r}
shape_select(
  x,
  text = "^\\s*\\{image_[0-9]+\\}\\s*$",
  match = "regex"
)[, c("slide_idx", "name", "text")]
```

## Filtering by Metadata

Different criteria are combined with AND. Multiple values within one criterion are alternatives.

```{r}
shape_select(
  x,
  text = "{image_1}",
  name = "marker"
)[, c("slide_idx", "shape_idx", "name", "text", "kind")]
```

Object kinds currently include:

| Kind | Meaning |
|------|---------|
| `"shape"` | Regular PowerPoint shape or text box. |
| `"picture"` | Picture object. |
| `"table"` | Table graphic frame. |
| `"chart"` | Chart graphic frame. |
| `"graphic_frame"` | Other graphic frame. |
| `"group"` | Top-level group. |
| `"connector"` | Connector shape. |

## Result Columns

`shape_select()` returns a tibble-like `pptx_shape_selection` object. The `node` list-column is kept in the object for downstream functions, but hidden by the print method.

Important columns:

| Column | Meaning |
|--------|---------|
| `slide_idx` | Visible presentation slide index. |
| `shape_idx` | Top-level shape-tree position, which approximates z-order. |
| `id` | Current slide-local XML shape ID. Not stable across save operations. |
| `name` | Selection Pane name. |
| `text` | Visible text. For groups this is `NA` in the first implementation. |
| `kind` | Object kind. |
| `placeholder` | Whether the object is an actual placeholder. |
| `ph_type` | Placeholder type, if present. |
| `hidden` | Whether `p:cNvPr/@hidden` is set to `"1"`. |
| `left`, `top`, `width`, `height` | Slide-level frame in inches. |
| `rotation` | Rotation in degrees, if present. |
| `description` | Alternative text description, if present. |
| `node` | XML node reference for downstream code. |

```{r}
targets <- shape_select(x, text = "{image_1}")
names(targets)
```

Selections also carry internal source metadata. Functions such as `shape_update()` use it to reject selections from a different presentation or selections that became stale after saving, reparsing, deleting shapes, or changing slide order. If that happens, rerun `shape_select()`.

## Updating Selected Shapes

Use `shape_update()` when a selected marker should be changed in place. The function returns the modified `rpptx` object.

```{r}
targets <- shape_select(x, text = "{image_1}", match = "exact")

x <- shape_update(
  x,
  targets,
  text = "Delayed",
  background = "#FCE8E6",
  line = sp_line(color = "#EC0016", lwd = 1.5),
  name = "Status marker",
  description = "Current service status",
  hidden = FALSE
)

shape_select(x, name = "Status marker", match = "exact")[
  , c("name", "text", "description", "hidden")
]
```

`shape_update()` edits existing XML nodes. It does not remove and recreate shapes, so z-order and shape identity are preserved.

Supported update arguments:

| Argument | Meaning |
|----------|---------|
| `text` | Replace the complete visible text. |
| `background` | Set a solid fill. Use `"transparent"` for no fill. |
| `line` | Apply an `officer::sp_line()` outline. |
| `left`, `top`, `width`, `height` | Update position and size in inches. |
| `rotation` | Update rotation in degrees. |
| `name` | Update the Selection Pane name. |
| `description` | Update alt text description. Use `NA` to remove it. |
| `hidden` | Hide or show the shape. |

Different shape kinds have different capabilities. Regular shapes support all update arguments. Connectors support line, geometry, metadata, and visibility. Pictures support line, geometry, metadata, and visibility. Graphic frames and groups currently support geometry, metadata, and visibility.

### Updating Multiple Shapes

`shape_update()` updates every row in the selection. Scalar values are recycled to all selected shapes. Vectors and lists with the same length as the selection are applied row by row.

The following example selects three marker shapes and updates their text, fill, outline, and alt-text descriptions in one call.

```{r}
status_deck <- function() {
  x <- read_pptx()
  x <- add_slide(x, layout = "Blank")
  x <- ph_with(
    x,
    "Operational status markers",
    location = ph_location(
      left = 0.6, top = 0.35, width = 6.5, height = 0.45,
      newlabel = "Demo title"
    )
  )

  for (i in 1:3) {
    x <- ph_with(
      x,
      paste0("{status_", i, "}"),
      location = ph_location(
        left = 0.8 + (i - 1) * 2.35,
        top = 1.45,
        width = 1.9,
        height = 1.05,
        newlabel = paste0("Status marker ", i),
        bg = "#F5F5F5",
        ln = sp_line(color = "#878C96", lwd = 1.2)
      )
    )
  }
  x
}

template <- status_deck()
targets <- shape_select(template, text = "{status_", match = "contains")

updated <- shape_update(
  template,
  targets,
  text = c("On time", "Delayed", "Cancelled"),
  background = c("#E6F4EA", "#FFF4CE", "#FCE8E6"),
  line = list(
    sp_line(color = "#408335", lwd = 1.5),
    sp_line(color = "#F39200", lwd = 1.5),
    sp_line(color = "#EC0016", lwd = 1.5)
  ),
  description = c(
    "Green status marker",
    "Yellow status marker",
    "Red status marker"
  )
)

shape_select(updated, name = "Status marker", match = "contains")[
  , c("name", "text", "description")
]
```

The thumbnails show the marker slide before and after the update:

```{r}
#| echo: false
#| out-width: "49%"
#| fig-show: hold

knitr::include_graphics(shape_thumbnail(c(
  "thumb-shape-update-before.png",
  "thumb-shape-update-after.png"
)))
```

This vignette creates the demo markers with `ph_with()` so the example is reproducible from an empty presentation. `shape_update()` itself does not depend on layout placeholders; it updates the selected slide XML nodes. The test suite covers ordinary non-placeholder shapes with `<p:nvPr/>` and no `<p:ph/>`.

## Groups

The first implementation is top-level only. A group is returned as one row with `kind = "group"`; shapes inside the group are not listed or matched.

This keeps coordinates simple and reliable. Recursive group support would need separate handling for group-relative coordinates, nested z-order, rotation, and possible duplicate matches between a group and its children.

## Relation to officer Placeholders

`shape_select()` works on actual slide XML nodes. It is different from officer's `ph_location_*()` functions, which resolve layout placeholders.

Use `ph_location_*()` when you are creating new content from a layout. Use `shape_select()` when an existing slide already contains marker shapes that should be discovered.

`shape_update()` follows the same rule. It updates selected slide XML nodes and does not require those nodes to be placeholders. Ordinary PowerPoint shapes with `<p:nvPr/>` and no `<p:ph/>` are supported and covered by tests.

## Future Extensions

Future functions can build on the same node-level setters to add new shapes, replace images, format text runs, reorder shapes, or edit shapes inside groups.

```{r}
#| eval: false
targets <- shape_select(template, text = "{image_1}")
template <- shape_update(template, targets, hidden = TRUE)
```
